---
title: "DSC week 14 IP part 3 and 4 with R"
author: "william"
date: "23/01/2021"
output: html_document
---

Part 3: Association Rules

This section will require that you create association rules that will allow you to identify relationships between variables in the dataset. You are provided with a separate dataset that comprises groups of items that will be associated with others. Just like in the other sections, you will also be required to provide insights for your analysis.



## installing the package to be used
```{r}
#install the required arules library 
#install.packages("arules")
#install.packages("data.table")

library(arules)
library(data.table)
library(lattice)
library(ggplot2)
library(caret)
#install.packages(c("Rcpp","tidyverse")) # install packages to work with data frame - extends into visualization
library(tidyverse)
```



##2.0 Importing dataset

We will use read.transactions fuction which will load data from comma-separated files 
and convert them to the class transactions, which is the kind of data that 
we will require while working with models of association rules

```{r}
# Dataset Url = http://bit.ly/SupermarketDatasetII
# Importing our dataset
# ---
#
super_dataset <- read.transactions('http://bit.ly/SupermarketDatasetII',sep = ",")

# Previewing the dataset
# ---
#having a look at the dataset dimension, and variables
dim(super_dataset)


```
observation: our dataset has 1701 records and 119 variables. 


### checking data type
```{r}
# Verifying the object's class
# ---
# This should show us transactions as the type of data that we will need
# ---
# 
class(super_dataset)
```
observation: out dataset has transaction data type which is the right one for this case.



```{r}
# Previewing our first 5 transactions
#
inspect(super_dataset[1:5])
```
observation: the above shows the first 5 transactions.


### previewing the dataset
```{r}
# If we wanted to preview the items that make up our super_dataset dataset,
# alternatively we can do the following
# ---
# 
items<-as.data.frame(itemLabels(super_dataset))
colnames(items) <- "Item"
head(items, 10) 
```

observation: the above shows the ten items that make up our dataset. 


### checking the dataset summary
```{r}
# Generating a summary of the super-dataset dataset
# ---
# This would give us some information such as the most purchased items, 
# distribution of the item sets (no. of items purchased in each transaction), etc.
# ---
# 
summary(super_dataset)
```
this is a concise summary of our dataset. it gives us the most frequent items, which includes mineral water, eggs, spaghetti, french fries, and chocolates.

It also gives us the minimum, median,mean and maximum values in our dataset.


### checking the items frequency
```{r}
# Exploring the frequency of some articles 
# of transactions ranging from 1 to 5 and performing 
# some operation in percentage terms of the total transactions 
# 
itemFrequency(super_dataset[, 1:5],type = "absolute")
round(itemFrequency(super_dataset[, 1:5],type = "relative")*100,2)
```
observation: we can avocado has the highest stock of 250 whic is 3.33% of the total transaction, followed by almond at 153 which is 2.04% of the total transaction, and so on.


### visualization
```{r}
# Producing a chart of frequencies and filtering 
# to consider only items with a minimum percentage 
# of support/ considering a top x of items
# ---
# Displaying top 10 most common items in the transactions dataset 
# and the items whose relative importance is at least 10%
# 
par(mfrow = c(1, 2))

# plot the frequency of items
itemFrequencyPlot(super_dataset, topN = 10,col="darkgreen")
itemFrequencyPlot(super_dataset, support = 0.1,col="darkred")
```
observation:

The most common items ara mineral water, eggs, spaghetti, french fries, chocolate, green tea, milk, ground beef, frozen vegetable and pancakes.

The left plot shows items whose relative importance is at least 10%, they include chocolates, eggs, french fries, green tea, milk, mineral water and spaghetti.



### Building the model using association rules and apriori function
```{r}
# Building a model based on association rules 
# using the apriori function 
# ---
# We use Min Support as 0.001 and confidence as 0.8
# ---
# 
rules1 <- apriori (super_dataset, parameter = list(supp = 0.001, conf = 0.8))
rules1

```

observation, we build the model using association rules and apriori() function, min support=0.001, confidence=0.8

we got 74 rules with this conditions.

However, in order to illustrate the sensitivity of the model to these two parameters, 
we will increase the support or lower the confidence level and see the result.



### increasing the support and lowering the confidence level
```{r}
# Building a apriori model with Min Support as 0.002 and confidence as 0.8.
rules2 <- apriori (super_dataset,parameter = list(supp = 0.002, conf = 0.8)) 

# Building apriori model with Min Support as 0.002 and confidence as 0.6.
rules3 <- apriori (super_dataset, parameter = list(supp = 0.001, conf = 0.6)) 

rules2

rules3
```

In rule 2, we increased the minimum support of 0.001 to 0.002 and model rules went from 72 to only 2. This would lead us to understand that using a high level of support can make the model lose interesting rules. 

In rule 3, we decreased the minimum confidence level to 0.6 and the number of model rules went from 72 to 545. This would mean that using a low confidence level increases the number of rules to quite an extent and many will not be useful.


### performing exploration of the model using summary function
```{r}
# We can perform an exploration of our model 
# through the use of the summary function as shown
# ---
# Upon running the code, the function would give us information about the model 
# i.e. the size of rules, depending on the items that contain these rules. 
# In our above case, most rules have 3 and 4 items though some rules do have upto 6. 
# More statistical information such as support, lift and confidence is also provided.
# ---
# 
summary(rules1)
```
observation: we are able to see the min, max, median, mean, and max of the rule 1 as well as rule length.


###checking the first 5 rules 
```{r}
# Observing rules built in our model i.e. first 5 model rules
# ---
# 
inspect(rules1[1:5])

```
observation:

 If someone buys frozen smoothie and spinach, they are 88% likely to buy mineral water too.
 
 If someone buys bacon and pancake, they are 81% likely to buy spaghetti too.
 
 If someone buys mushroom cream sauce, and pasta, they are 95% likely to buy escalope too etc.
  


### ordering these rules by criteria
```{r}
# Ordering these rules by a criteria such as the level of confidence
# then looking at the first five rules.
# We can also use different criteria such as: (by = "lift" or by = "support")
# 
rules1<-sort(rules1, by="confidence", decreasing=TRUE)
inspect(rules1[1:5])

```
Interpretation:

The given five rules have a confidence of 100, this means

 If someone buys rice, and sugar, they are 100% likely to buy whole milk too.
 
 If someone buys canned fish, and hygiene articles, they are 100% likely to buy whole milk too.
 

###. visualizing rules1
```{r}
library(arulesViz)
plot(rules1,method = "graph",)
```

The above shows the top 72 rules ordered by support


### making promotion
```{r}
# If we're interested in making a promotion relating to the sale of spaghetti, 
# we could create a subset of rules concerning these products 
# ---
# This would tell us the items that the customers bought before purchasing spaghetti
# ---
# 
spaghetti<- subset(rules1, subset = rhs %pin% "spaghetti")
 
# Then order by confidence
spaghetti<-sort(spaghetti, by="confidence", decreasing=TRUE)

inspect(spaghetti[1:5])

```
explanation:

these are the items that will be bought before buying spaghetti.



###. checking the top 5 items that will be bought before buying spagetti

```{r}
# What if we wanted to determine items that customers might buy 
# who have previously bought yogurt?
# ---
# 
# Subset the rules
spaghttie <- subset(rules1, subset = lhs %pin% "spaghetti")

# Order by confidence
spaghetti<-sort(spaghetti, by="confidence", decreasing=TRUE)

```

```{r}
# inspect top 5
inspect(spaghetti[1:5])
```
explanation:

the customer is likely to buy these top 5 items  before he can buy spaghetti.


```{r}
plot(spaghetti,method = "graph",)
```
The above shows the top 10 rules ordered by support; it shows the items that are liked to be bought along with spaghetti. 


## checking the optimal combinations

### installing the package to be used
```{r}
library("devtools")
#install_github("mhahsler/arulesViz")1
library(arulesViz)
```


### using apriopi function to build the rules
```{r}
##using apriori function to build the rules
r<- apriori (super_dataset, parameter = list(supp = 0.001, conf = 0.9, minlen=4))

```


### getting the summary
```{r}
summary(r)
```
the above shows the summary of our dataset r


### visualizing the rules
```{r}
# Visualising our rules
plot(r,method = "scatterplot",jitter=0)
```
observation: From the visualization we can conclude that our rules have a good lift , confidence and support.


### ploting the graph
```{r}
plot(r,method = "graph",)
```
The above shows the top 100 rules ordered by support and the items  can be bought together.



### Part 4: Anomaly Detection

You have also been requested to check whether there are any anomalies in the given sales dataset. The objective of this task being fraud detection.


```{r}
# Dataset Url = http://bit.ly/CarreFourSalesDataset
# Importing our dataset
# ---
#
carrefour_dataset <- read.csv('http://bit.ly/CarreFourSalesDataset')

# Previewing the dataset
# ---
#having a look at the dataset dimension, and variables
dim(carrefour_dataset)
head(carrefour_dataset)


```
observation: the dataset has 1000 records and 2 variables. 


### installing necessary packages

```{r}
# Install Lubridate package for date time data
#install.packages('lubridate')
library(lubridate)
```


```{r}
# Installing anomalize package
#
#install.packages("anomalize")
#
# Load tidyverse and anomalize 
library(tidyverse)
library(anomalize)
```


### changing date to date time data
```{r}
#changing date to date time. 
carrefour_dataset$Date <- as.Date(carrefour_dataset$Date, "%m/%d/%y")
head(carrefour_dataset)
```


### Detecting our anomalies



### checking  the number of transaction in a day

```{r}
# group and tally the number of transactions per day
carrefour_df <- carrefour_dataset %>% group_by(Date) %>% tally()
colnames(carrefour_df) <- c('transactionDate', 'totalCount')
head(carrefour_df)
```
the above shows the number of the customer transactions frm Jan 1st to Jan 6th. 


### checking for anomalies and ploting them
```{r}
# anomalize() - 
# We perform anomaly detection on the decomposed data using 
# the remainder column through the use of the anomalize() function 
# which procides 3 new columns; “remainder_l1” (lower limit), 
# “remainder_l2” (upper limit), and “anomaly” (Yes/No Flag).
# The default method is method = "iqr", which is fast and relatively 
# accurate at detecting anomalies. 
# The alpha parameter is by default set to alpha = 0.05, 
# but can be adjusted to increase or decrease the height of the anomaly bands, 
# making it more difficult or less difficult for data to be anomalous. 
# The max_anoms parameter is by default set to a maximum of max_anoms = 0.2 
# for 20% of data that can be anomalous. 
# 
```


## ploting to visualize the anormaly
```{r}
# we now plot using plot_anomaly_decomposition() to visualize out data.
carrefour_df %>%
    time_decompose(totalCount) %>%
    anomalize(remainder) %>%
    plot_anomaly_decomposition(ncol = 2, alpha_dots = 0.5) +
    ggtitle("Anomaly Detection Plot")
```
observation: there is no red flag thus these customer transactions do not have anomalies.



### plotting recomposition to check anomalies further
```{r}
# ploting the recomposition to try and see anomalies
#
carrefour_df %>%
    time_decompose(totalCount) %>%
    anomalize(remainder) %>%
    time_recompose() %>%
    plot_anomalies(time_recomposed = TRUE, ncol = 4, alpha_dots = 0.25, fill="dark red") +
    ggtitle("Anomalie detection plots")

```

observation: we can conclude that there is no anomaly(outliers) in this dataset.
